On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-06 manuscripts/PNAS/chazthm_pnas.tex
main lemma

\begin{proof}[Proof of Main Lemma]
We first claim $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$ for all $S \in E$. To see why, note that the right-hand side of \eqref{GapUpperBound} is then strictly less than one, since $r \leq 1$ and $L_2(\mathbf{A}) \leq 1$ by \eqref{delrho}. By \eqref{dimLem}, it then follows that $|\pi(S)| \geq \dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \geq \dim(\text{\rm Span}\{\mathbf{A}_{S}\}) = |S|$, with the equality due to $\mathbf{A}$ satisfying \eqref{SparkCondition}; hence, $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$ (since $|S| = |\pi(S)|$). It follows that the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent and, by \eqref{eqdim}:% and $\mathbf{B}_i \neq \textbf{0}$ for all $i$.
\begin{align}\label{eq2}
d(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_S\} ) = d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}).
\end{align}
%We will now show that \eqref{cond} holds.
%\begin{align}\label{fact2}
%|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } \ F \subseteq E.
%\end{align}

Fix $F \subseteq E$. Since $\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\} \subseteq \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}$, if $\cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\} = \textbf{0}$ then we must have $|\cap_{S \in F} \pi(S)| = 0$ (as $\mathbf{B}_i \neq \textbf{0}$ for all $i$) and \eqref{cond} trivially holds. Suppose then that the intersection is not the zero vector. By Lem.~\ref{SpanIntersectionLemma} and Lem.~\ref{DistanceToIntersectionLemma}, and incorporating \eqref{eq2} and \eqref{GapUpperBound}, we have:
\begin{align}\label{randoml}
d( \text{\rm Span}&\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\}  ) \nonumber \\
&\leq d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \cap_{S \in F} \text{\rm Span}\{\mathbf{A}_{S}\} \right) \nonumber \\
&\leq \sum_{S \in F} \frac{ d\left( \cap_{T \in F} \text{\rm Span}\{\mathbf{B}_{\pi(T)}\},\text{\rm Span}\{\mathbf{A}_{S}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{T}\}_{T \in F}) } \nonumber \\
&\leq \sum_{S \in F} \frac{ d\left( \text{\rm Span}\{\mathbf{B}_{\pi(S)}\},\text{\rm Span}\{\mathbf{A}_{S}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{T}\}_{T \in F}) }\nonumber \\
%&\leq \frac{1}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F})} \sum_{S \in F} \frac{ \varepsilon }{ L(\mathbf{AX}_{J(S)}) } \nonumber \\
&\leq \frac{|F| \varepsilon}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F})} 
%&< L_2(\mathbf{A}) \frac{|F| }{ 2^{|E|}} \left( \frac{ \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) } \right) \left( \frac{\min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }{ \min_{T \in F} L(\mathbf{AX}_{J(T)}) } \right) \nonumber \\
\leq C_2 \varepsilon. 
\end{align}
Note that since $\varepsilon < L_2(\mathbf{A}) / \tilde C_2$, by \eqref{delrho} the right-hand side in \eqref{randoml} is strictly less than one, so \eqref{dimLem} implies that $\dim(\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}) \leq \dim(\text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\})$ and \eqref{cond} follows from the linear independence of the columns of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in F$.
Now, fix $\ell \in [m]$. Since $E$ satisfies the singleton intersection property, $\{\ell\} = \cap F(\ell)$ for the star $F(\ell) \subseteq E$. By \eqref{cond}, we have that $\cap_{S \in F(\ell)} \pi(S)$ is either empty or it contains a single element. Lem.~\ref{NonEmptyLemma} ensures the latter case is the only possibility. Thus, the association $\ell \mapsto \cap_{S \in F(\ell)} \pi(S)$ defines a map $\bar \pi: [m] \to [m]$ and $\dim(\text{\rm Span}\{\mathbf{B}_{\bar \pi(\ell)}\}) = 1$, since $\mathbf{B}_i \neq \textbf{0}$ for all $i$. By \eqref{eqdim}, it follows from \eqref{randoml} that $d\left( \text{\rm Span}\{\mathbf{A}_\ell\}, \text{\rm Span}\{ \mathbf{B}_{\bar \pi(\ell)} \} \right) \leq C_2 \varepsilon$. Since $\ell$ is arbitrary, fixing $\bar \varepsilon = C_2\delta$ it follows from \eqref{randoml} that for every basis vector $\mathbf{e}_\ell \in \mathbb{R}^m$ there exists some $c'_\ell \in \mathbb{R}$ such that $\|\mathbf{A}\mathbf{e}_\ell - c'_\ell \mathbf{B}\mathbf{e}_{\bar \pi(\ell)}\|_2 \leq \bar \varepsilon < L_2(\mathbf{A})$. This is exactly the supposition in \eqref{1D} (letting $c_i = \|\mathbf{A}_i\|_2^{-1}$ for all $i$) and the result follows from the subsequent arguments for the case $k=1$ in Sec.~\ref{DUT}.
\end{proof}

>>>>

\begin{proof}[Proof of Main Lemma]
We begin by showing that: 
\begin{align}\label{eq2}
d(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_S\} ) = d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}).
\end{align}
%We first claim $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$ for all $S \in E$. 
Since the right-hand side of \eqref{GapUpperBound} is strictly less than one (by \eqref{delrho} and $r \leq 1$), it follows by \eqref{dimLem} that $|\pi(S)| \geq \dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \geq \dim(\text{\rm Span}\{\mathbf{A}_{S}\}) = |S|$, with the equality due to $\mathbf{A}$ satisfying \eqref{SparkCondition}. Since $|S| = |\pi(S)|$, we therefore have $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$, and \eqref{eq2} follows by \eqref{eqdim}. Note that the columns of $\mathbf{B}_{\pi(S)}$ are therefore linearly independent, for all $S \in E$.

We will now show that \eqref{cond} holds.
%\begin{align}\label{fact2}
%|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for all } \ F \subseteq E.
%\end{align}
Fix $F \subseteq E$. Since $\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\} \subseteq \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}$, if $\cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\} = \textbf{0}$ then we must have $|\cap_{S \in F} \pi(S)| = 0$ (as the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent) and \eqref{cond} trivially holds. Suppose then that the intersection is not the zero vector. By Lem.~\ref{SpanIntersectionLemma} and Lem.~\ref{DistanceToIntersectionLemma}, and incorporating \eqref{eq2} and \eqref{GapUpperBound}, we have:
\begin{align}\label{randoml}
d( \text{\rm Span}&\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\}  ) \nonumber \\
&\leq d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \cap_{S \in F} \text{\rm Span}\{\mathbf{A}_{S}\} \right) \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) } \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \text{\rm Span}\{\mathbf{B}_{\pi(T)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) }\nonumber \\
%&\leq \frac{1}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F})} \sum_{S \in F} \frac{ \varepsilon }{ L(\mathbf{AX}_{J(S)}) } \nonumber \\
&\leq \frac{|F| \varepsilon}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F})} 
%&< L_2(\mathbf{A}) \frac{|F| }{ 2^{|E|}} \left( \frac{ \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) } \right) \left( \frac{\min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }{ \min_{T \in F} L(\mathbf{AX}_{J(T)}) } \right) \nonumber \\
\leq \frac{C_2 \varepsilon}{\max_i\|\mathbf{A}_i\|_2}. 
\end{align}
Since $\varepsilon < L_2(\mathbf{A}) / C_2$, by \eqref{delrho} the right-hand side in \eqref{randoml} is strictly less than one. Hence, $\dim(\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}) \leq \dim(\text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\})$ by \eqref{dimLem}, and \eqref{cond} follows from the linear independence of the columns of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in F$.

%Now, fix $\ell \in [m]$. Since $E$ satisfies the singleton intersection property, $\{\ell\} = \cap F(\ell)$ for the star $F(\ell) \subseteq E$. By \eqref{cond}, we have that $\cap_{S \in F(\ell)} \pi(S)$ is either empty or it contains a single element. Lem.~\ref{NonEmptyLemma} ensures the latter case is the only possibility. 
Now, by Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap \pi(F(i))$ defines an injective map $\bar \pi: J \to [\bar m]$ for some $J \subseteq [m]$ of size $\bar m - \ell(\bar m - m)$, and we can be sure that $\mathbf{B}_{\bar \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in E$. It then follows from \eqref{eqdim} and \eqref{randoml} that $d\left( \text{\rm Span}\{\mathbf{A}_i\}, \text{\rm Span}\{ \mathbf{B}_{\bar \pi(i)} \} \right) \leq C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$ for all $i \in J$. Fixing $\bar \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for every basis vector $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $c'_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - c'_i \mathbf{B}\mathbf{e}_{\bar \pi(i)}\|_2 \leq \bar \varepsilon < L_2(\mathbf{A}) \min_{i\in J} c_i$. This is exactly the supposition in \eqref{1D}, and the result follows from the subsequent arguments for the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$.
\end{proof}
